---
title: "Green hydrogen upscaling"
author: "Adrian Odenweller"
output:
  html_document:
    df_print: paged
---

Main analysis file for the  article:
*Probabilistic feasibility space of scaling up green hydrogen supply*

# Setup

```{r setup, warning = FALSE}
library(tidyverse)  # Data handling
library(quitte)  # Data handling
library(readxl)  # Reading excel files
library(ggplot2)  # Plotting
library(cowplot)  # Plotting layout
library(ggnewscale)  # Multiple legends
library(ggExtra)  # Marginal distributions
library(ggsci)  # Colour palettes
library(scales)  # Plotting axes
library(zoo)  # Function na.approx
library(ggforce)  # Function facet_zoom
library(colorspace)  # Colour manipulation
library(nleqslv)  # Solving non-linear systems of equations
library(truncnorm)  # Truncated normal distribution
library(lubridate)  # Working with dates

# Input data folder
path.input <- "01_input_data/"
# Output folder
path.output <- "02_output_plots/"
# Helper functions folder
path.helpers <- "03_helpers/"
# Plotting functions folder
path.plotting <- "04_plotting/"

# Source helper functions
source(paste0(path.helpers, "ReadIEAHydrogenProjectsDB.R"))
source(paste0(path.helpers, "ReadBPDB.R"))
source(paste0(path.helpers, "calcTruncNormParam.R"))
source(paste0(path.helpers, "calcMonteCarloSample.R"))
source(paste0(path.helpers, "calcTechnologyDiffusion.R"))
source(paste0(path.helpers, "calcGompertzParam.R"))
source(paste0(path.helpers, "calcGompertzModel.R"))

# Source plotting functions
source(paste0(path.plotting, "plotElectrolysisProjects.R"))
source(paste0(path.plotting, "plotIllustration.R"))
source(paste0(path.plotting, "plotParameterDistributions.R"))
source(paste0(path.plotting, "plotProbabilisticFeasibilitySpace.R"))
source(paste0(path.plotting, "plotComparisonBreakthrough.R"))
source(paste0(path.plotting, "plotComparisonCapacity.R"))
source(paste0(path.plotting, "plotWindSolar.R"))

# Mode of model execution
#  Reproduction mode: Download pre-run simulation output from Zenodo first
#  Simulation mode: 
model.mode <- "reproduction"
#model.mode <- "simulation"

# Sample size (change to decrease computational effort)
n <- 10000
# Minimum annual growth rate
bmin <- 0.15

# Modelling assumptions and conversion factors
# Full load hours
flh <- 5000
# 60 % total efficiency on final energy level due to mix of hydrogen/e-fuels and transport losses
eff.fe <- 0.6
# 70 % efficiency on secondary energy level
eff.se <- 0.7
ej2gwh <- 10^6/3.6  # EJ to GWh
ej2gw <- ej2gwh / (flh*eff.se)  # EJ to GW

# For plotting
t.split <- 2023
t.min <- 2000
t.max <- 2030

# Set plotting style
font.size <- 7
theme_set(theme_cowplot(font_size = font.size))
color.pv <- "#ffb200"
color.wind <- "#337fff"

# Quantile calculation
quibble <- function(x, q = c(0.25, 0.33, 0.5, 0.66, 0.75)) {
  tibble(x = quantile(x, q), q = q)
}

# Regions
region.mapping.eu <- c(
  "Denmark" = "EU",
  "France" = "EU",
  "Germany" = "EU",
  "Netherlands" = "EU",
  "Spain" = "EU",
  "Rest EU" = "EU"
)
region.mapping.global <- c(
  "Denmark" = "EU",
  "France" = "EU",
  "Germany" = "EU",
  "Netherlands" = "EU",
  "Spain" = "EU",
  "Rest EU" = "EU",
  "MENA" = "MENA",
  "N. America" = "N. America",
  "C. + S. America" = "C. + S. America",
  "Asia" = "Asia",
  "Australia" = "Australia",
  "Other" = "Other"
  )
```

## Check file availability

```{r}
if (model.mode == "reproduction"){
  print("Model started in reproduction mode.")
  files.required <- c("02_output_plots/01_conventional_growth_parameters.rds",
                      "02_output_plots/02_conventional_growth_sample.rds",
                      "02_output_plots/03_conventional_growth_results.rds",
                      "02_output_plots/04_unconventional_growth_parameters.rds",
                      "02_output_plots/05_unconventional_growth_sample.rds",
                      "02_output_plots/06_unconventional_growth_results.rds")
  if (all(file.exists(files.required)) == TRUE){
    print("All required simulation result files found.")
  } else
    print("Required simulation result files are missing.")
} else if (model.mode == "simulation"){
  print("Model started in simulation mode.")
  } else {
  stop("Please specify either reproduction or simulation mode.")
}
```

# IEA Hydrogen Projects Database

```{r, message = FALSE}
# IEA Hydrogen Projects Database
file.h2projects <- paste0(path.input, "IEA Hydrogen Projects Database 2021 (revised).xlsx")
# Changes to IEA from own research
file.h2projects.changed <- paste0(path.input, "IEA Hydrogen Projects Database_CHANGES_ONLY.xlsx")
data.h2 <- ReadIEAHydrogenProjectsDB(
  file.h2projects,
  h2projects.range = "A5:AF994",  # ADJUST IF NECESSARY
  file.h2projects.changed,
  h2projects.changed.range = "A5:AF54"  # ADJUST IF NECESSARY
)
```

## FIGURE 1: Electrolysis projects

```{r, message = FALSE}
regions.h2 = tibble(name = c("EU", "Global"),
                    list = list(
                      names(region.mapping.eu),
                      unique(data.h2$region)
                    ))

# EU
regions <- regions.h2$list[[1]]
names(regions) <- regions

p1 <- plotElectrolysisProjects(data.h2 = data.h2,
                   regions = regions,
                   fill.by = "region",
                   title = "EU by country",
                   leg.label = "Country")

p2 <- plotElectrolysisProjects(data.h2 = data.h2,
                   regions = regions,
                   title = "EU by status",
                   fill.by = "status",
                   leg.label = "Status")

# Global
regions <- region.mapping.global

p3 <- plotElectrolysisProjects(data.h2 = data.h2,
                   regions = regions,
                   fill.by = "region",
                   title = "Global by region",
                   leg.label = "Region")

p4 <- plotElectrolysisProjects(data.h2 = data.h2,
                   regions = regions,
                   fill.by = "status",
                   title = "Global by status",
                   leg.label = "Status")

p <- plot_grid(p1, p3, p2, p4,
               ncol = 2,
               labels = "auto",
               label_size = font.size)

print(p)
ggsave(paste0(path.output, "FIG01_electrolysis_projects.png"),
       bg = "white", width = 18, height = 12, units = "cm")
ggsave(paste0(path.output, "FIG01_electrolysis_projects.pdf"),
       bg = "white", width = 18, height = 12, units = "cm")
```

# Conventional growth

## FIGURE 2: Illustration

```{r, message = FALSE, warning = FALSE}
p <- plotIllustration()

print(p)
ggsave(paste0(path.output, "FIG02_illustration.png"),
       width = 18, height = 12, units = "cm", bg = "white")
ggsave(paste0(path.output, "FIG02_illustration.pdf"),
       width = 18, height = 12, units = "cm", bg = "white")
```

## Input data

### Wind and solar power capacities

```{r, message = FALSE}
# BP Statistical Review of World Energy 2021
file.bp <- paste0(path.input, "bp-stats-review-2021-all-data.xlsx")
data.bp <- ReadBPDB(file.bp)
```

### Electrolysis deployment targets

```{r}
data.targets <- tribble(
  ~ region, ~year, ~unit, ~value,
  # EU hydrogen strategy
  "EU", 2024, "GW", 6,  # EU Fit for 55
  "EU", 2030, "GW", 100,  # REPowerEU: 10 Mt, approx. equiv. to 100 GW
  "EU", 2050, "GW", 500,  # EU Fit for 55
  # Global: IEA Hydrogen Projects Database and IEA NZE scenario
  "Global", 2030, "GW", 254,  # Cumulative announcements + 10 Mt EU imports
  "Global", 2050, "GW", 3600,  # IEA NZE
)
```


## Emergence growth rate from wind and solar power

```{r, message = FALSE}
# Fit exponential growth of wind and solar power in slices of varying lengths

# Region mapping
regions.bp = tibble(name = c("EU", "Global"),
                 list = list("EU",
                             "Global"),
                 span = list(1995:2010,
                             1995:2010))

time.slice.length <- 5:10

# Additive error specification
exp_fit <- function(df)
  nls(capacity ~ a*(1 + b)**(year-1995),
      data = df,
      start = list(a = 100,
                   b = 0.4))

# Get growth rate
get_growthrate <- function(mod)
  coef(mod)[["b"]]

# Get parameter a
get_a <- function(mod)
  coef(mod)[["a"]]

# Get predicted values
get_predict <- function(mod)
  list(predict(mod))

# Loop over regions and time slices
data.bp.fit <- NULL
# Loop over regions
for (r in 1:dim(regions.bp)[1]) {
  time.span <- regions.bp$span[[r]]
  # Loop over lengths of time slice
  for (t in time.slice.length){
    # Get individual time slices for each length t
    for (s in 1:(length(time.span)-t+1)){
      slice <- time.span[s:(s+t-1)]
      
      temp <- data.bp %>% 
        filter(region %in% regions.bp$list[[r]],
               year %in% slice) %>% 
        group_by(technology, year) %>% 
        summarise(capacity = sum(capacity)) %>% 
        group_by(technology) %>% 
        nest() %>% 
        # Fit exponential models using purrr::map
        mutate(model.exp = map(data, exp_fit)) %>% 
        transmute(technology,
                  b = map_dbl(model.exp, get_growthrate),
                  a = map_dbl(model.exp, get_a),
                  predict = map(model.exp, get_predict),
                  slice.length = t,
                  slice.start = slice[1],
                  region = regions.bp$name[[r]])
     
      data.bp.fit <- bind_rows(data.bp.fit, temp)
    }
  }
}
```

## Monte Carlo sampling: Initial capacity, emergence growth rate

```{r, message = FALSE}
# Region mapping
regions.h2 = tibble(name = c("EU", "Global"),
                    list = list(
                      names(region.mapping.eu),
                      unique(data.h2$region)
                    ))

if (model.mode == "simulation") {
  temp <- calcMonteCarloSample(
    # Sample size
    N = n,
    # Initial value distribution
    a.status = c("Decommissioned", "Operational", "Under construction"),
    # Lower truncation
    cap.1 = c("Decommissioned", "Operational", "Under construction", "FID"),
    # Condition 1
    cumprob.1 = 0.15,
    # Cumulative probability of condition 1
    feasibility.success.rate = 0.3,
    # Condition 2: x % of feasibility study projects
    # Growth rate distribution
    slice.len = 7,
    # Length of time slice (wind, solar)
    b.min = bmin  # Lower truncation
  )
  data.input <- temp[[1]]
  data.input.param <- temp[[2]]
  saveRDS(data.input.param,
          "02_output_plots/01_conventional_growth_parameters.rds")
  saveRDS(data.input,
          "02_output_plots/02_conventional_growth_sample.rds")
} else if (model.mode == "reproduction") {
  data.input.param <-
    readRDS("02_output_plots/01_conventional_growth_parameters.rds")
  data.input <-
    readRDS("02_output_plots/02_conventional_growth_sample.rds")
}
```

## FIGURE 3: Parameter distributions

```{r, warning = FALSE}
plots <- plotParameterDistributions(data.input = data.input,
                                    data.input.param = data.input.param)

# Total plot
p.col1 <- plot_grid(plots[[1]] + theme(legend.position = "none"),
                    plots[[2]] + theme(legend.position = "none"),
                    NULL,
                    plots[[3]] + theme(legend.position = "none"),
                    ncol = 1,
                    align = "v",
                    rel_heights = c(1, 0.15, 0.05, 1.1),
                    labels = c("a", "", "","c"),
                    label_size = font.size
                    )

p.col2 <- plot_grid(plots[[4]] + theme(legend.position = "none"),
                    plots[[5]] + theme(legend.position = "none"),
                    NULL,
                    plots[[6]] + theme(legend.position = "none"),
                    ncol = 1,
                    align = "v",
                    rel_heights = c(1, 0.15, 0.05, 1.1),
                    labels = c("b", "", "","d"),
                    label_size = font.size
                    )

p.plots <- plot_grid(p.col1,
                     p.col2,
                     nrow = 1)

p.legend <- plot_grid(NULL,
                      get_legend(plots[[1]] + theme(legend.position = "bottom")),
                      get_legend(plots[[2]] + theme(legend.direction = "vertical")),
                      NULL,
                      nrow = 1,
                      rel_widths = c(1,0.9,0.7,1))

p <- plot_grid(p.plots,
               p.legend,
               ncol = 1,
               rel_heights = c(1,0.15))

print(p)
ggsave(paste0(path.output, "FIG03_parameter_distributions.png"),
       width = 18, height = 14, units = "cm", bg = "white")
ggsave(paste0(path.output, "FIG03_parameter_distributions.pdf"),
       width = 18, height = 14, units = "cm", bg = "white")
```

## Demand pull

```{r}
# Different anticipation scenarios (in years)
anticipation <- c(0, 5, 10, 100)  # 100y = full demand pull

# Model time span
model.time <- t.split:2050

data.demand <- NULL
for (r in 1:2){
  for (a in anticipation){
    
    # Targets in region
    targets.temp <- data.targets %>%
      filter(region == regions.h2$name[[r]])
    
    # Calculate linear demand pull with anticipation
    data.temp <- tibble(region = regions.bp$name[r],
                        year = model.time) %>% 
      mutate(demand = approx(x = targets.temp$year,
                             y = targets.temp$value,
                             xout = model.time,
                             method = "linear",
                             rule = 2)[["y"]]) %>% 
      mutate(demand = lead(demand, a, default = last(demand)),
             anticipation = a) %>% 
      group_by(region) %>%
      nest(demand = c("year", "demand")) %>% 
      ungroup()
    
    data.demand <- bind_rows(data.demand, data.temp)
  }
}
```

## Call simulation

```{r, message = F}
# Simulation mode
if (model.mode == "simulation") {
  # Keep track of time
  start.time <- Sys.time()
  # Call diffusion model
  data.results <-
    calcTechnologyDiffusion(data.demand, data.input, 0.25)
  # Save simulation results
  saveRDS(data.results,
          "02_output_plots/03_conventional_growth_results.rds")
  # Print duration of simulation
  end.time <- Sys.time()
  time.taken <- end.time - start.time
  print(time.taken)
  # Reproduction mode
} else if (model.mode == "reproduction") {
  data.results <-
    readRDS("02_output_plots/03_conventional_growth_results.rds")
}
```

## FIGURE 4: Probabilistic feasibility space

```{r}
colours <- pal_npg()(2)

p1 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results %>%
    filter(region == "EU", anticipation == 5),
  data.targets = data.targets,
  colour = colours[1],
  marginal.year = 2038,
  title = "EU"
)

p2 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results %>%
    filter(region == "Global", anticipation == 5),
  data.targets = data.targets,
  colour = colours[1],
  marginal.year = 2045,
  title = "Global"
)

p.legend <- p1[[2]]

p.plots <- plot_grid(p1[[1]], NULL, p2[[1]],
                     ncol = 3,
                     rel_widths = c(1, 0.05, 1),
                     labels = c("a", "", "b"),
                     label_size = font.size)

p <- plot_grid(p.plots,
               NULL,
               p.legend,
               ncol = 3,
               rel_widths = c(1, 0.05, 0.15))

print(p)
ggsave(paste0(path.output, "FIG04_probabilistic_feasibility_space.png"),
       width = 18, height = 9, units = "cm", bg = "white")
ggsave(paste0(path.output, "FIG04_probabilistic_feasibility_space.pdf"),
       width = 18, height = 9, units = "cm", bg = "white")
```

## Quantiles of results

```{r, message = FALSE}
for (r in c("EU", "Global")){
  
  stats <- data.results %>% 
  filter(region == r,
         anticipation == 5) %>% 
  select(!demand) %>% 
  unnest(sensitivities) %>% 
  unnest(results) %>% 
  filter(year %in% c(2030, 2040, 2050)) %>% 
  group_by(year) %>% 
  summarise(quibble(forecast))
  
  print(stats)
}
```


# Unconventional growth

## Input data

### WW2 US aircraft

```{r}
#www.aia-aerospace.org/wp-content/uploads/2016/06/Aviation-Facts-and-Figures-1945.pdf

data.ww2aircraft <- tibble(
  year = 1935:1944,
  prod = c(459, 1141, 949, 1800, 2141, 6086, 19290, 47873, 85946, 96369)
) %>% 
  mutate(value = cumsum(prod),
         unit = "units",
         tech = "WW2 US aircraft") %>% 
  select(!prod) %>% 
  mutate(number = 1)

```

### WW2 US liberty ships

```{r}
file.libertyships <- paste0(path.input, "Liberty_ships.xlsx")

data.libertyships <- read_excel(file.libertyships, sheet = "matlab") %>% 
  select(year, LS_cum) %>% 
  rename(value = LS_cum) %>% 
  group_by(year) %>% 
  slice(n()) %>% 
  mutate(unit = "units",
         tech = "WW2 US liberty ships") %>%
  ungroup() %>% 
  add_row(tibble(year = 1938,
                 unit = "units",
                 tech = "WW2 US liberty ships",
                 value = 0)) %>%
  arrange(year) %>% 
  mutate(number = 2)
```

### US nuclear weapons

```{r}
#https://doi.org/10.2968%2F066004008

file.nuclearweapons <- paste0(path.input, "Nuclear_weapons.xlsx")

data.usnuclearweapons <- read_excel(file.nuclearweapons, range = "A1:B57") %>% 
  rename(year = Year,
         value = US) %>% 
  mutate(unit = "units",
         tech = "US Nuclear weapons") %>% 
  filter(year <= 1967) %>% 
  mutate(number = 3)
```

### Soviet nuclear weapons

```{r}
#https://doi.org/10.2968%2F066004008

data.sovietnuclearweapons <-
  read_excel(file.nuclearweapons, range = "A1:C57") %>%
  rename(year = Year,
         value = Russia) %>% 
  select(!US) %>% 
  mutate(unit = "units",
         tech = "Soviet Nuclear weapons") %>% 
  filter(!is.na(value), year <= 1986) %>% 
  mutate(number = 4)
```

### Nuclear power in France

```{r, warning = FALSE}
data.nuclear <-
  read_excel(file.bp,
             sheet = "Nuclear Generation - TWh",
             range = "B3:BE32",
             col_types = "numeric") %>%
  slice(n()) %>% 
  pivot_longer(cols = "1965":"2020", names_to = "year", values_to = "value") %>% 
  mutate(year = as.integer(year),
         value = as.double(value),
         unit = "TWh/yr",
         tech = "Nuclear France") %>% 
  filter(year <= 2005) %>% 
  mutate(number = 5)
```

### Internet host count

```{r}
file.internet <- paste0(path.input, "Internet_host_count.xlsx")

data.internet <- read_excel(file.internet) %>% 
  mutate(year = decimal_date(date),
         value = value/1E6,
         unit = "million hosts",
         tech = "Internet host count") %>% 
  select(year, value, unit, tech) %>% 
  arrange(year) %>% 
  filter(year < 2018) %>% 
  mutate(number = 6)
```

### China high-speed rail network

```{r}
#https://uic.org/passenger/highspeed/article/high-speed-database-maps

file.chinahsr <- paste0(path.input, "china_hsr.xlsx")

data.chinahsr <- read_excel(file.chinahsr, sheet = "values") %>% 
  group_by(year) %>% 
  summarise(new = sum(km)) %>% 
  mutate(value = cumsum(new),
         unit = "km",
         tech = "High-speed rail China") %>% 
  select(year, value, unit, tech) %>% 
  complete(year = seq(2003, 2020)) %>% 
  fill(value, unit, tech) %>% 
  add_row(tibble(year = 2002,
                 value = 0,
                 unit = "km",
                 tech = "High-speed rail China")) %>% 
  arrange(year) %>% 
  mutate(number = 7)
```

### US dry shale gas

```{r}
#https://www.eia.gov/dnav/ng/hist/res_epg0_r5302_nus_bcfa.htm

file.shale <- paste0(path.input, "RES_EPG0_R5302_NUS_BCFa.xls")

data.shale <- read_excel(file.shale,
                         sheet = "Data 1",
                         range = "A4:B16",
                         col_names = c("date", "value")) %>% 
  mutate(unit = "billion cubic feet",
         tech = "US shale gas",
         year = decimal_date(date)) %>% 
  select(year, value, unit, tech) %>% 
  mutate(number = 8)
```

### Smartphone shipments (cumulative)

```{r}
#https://www.statista.com/statistics/263441/global-smartphone-shipments-forecast/
#https://www.statista.com/statistics/271539/worldwide-shipments-of-leading-smartphone-vendors-since-2007/
#https://www.researchgate.net/publication/50836092_Structuring_the_Smartphone_Industry_Is_the_Mobile_Internet_OS_Platform_the_Key/figures

data.smartphones <- tibble(
  year = 2006:2020,
  prod = c(64.1, 122.36, 151.4, 173.5, 304.7, 494.5, 725.3, 1018.7, 1301.7, 1437.2, 1473.4, 1465.5, 1402.6, 1371.1, 1280)
) %>% 
  mutate(value = cumsum(prod),
         unit = "million units",
         tech = "Smartphones globally") %>% 
  select(!prod) %>% 
  mutate(number = 9)
```

### Europe e-bike sales (cumulative)

```{r}
#https://www.statista.com/statistics/276036/unit-sales-e-bikes-europe/
#https://www.ziv-zweirad.de/fileadmin/redakteure/Downloads/Marktdaten/PM_2021_10.03._ZIV-Praesentation_10.03.2021_mit_Text.pdf

data.ebikes <- tibble(
  year = 2009:2020,
  sales = c(0.5, 0.7, 0.9, 1.1, 1, 1.25, 1.5, 1.74, 2.2, 2.92, 3.6, 5.1)
) %>% 
  mutate(value = cumsum(sales),
         unit = "million units",
         tech = "E-Bikes Europe") %>% 
  select(!sales) %>% 
  mutate(number = 10)
```

### COVID-19 vaccinations

```{r}
#https://ourworldindata.org/covid-vaccinations

file.covid <- paste0(path.input, "owid_vaccinations.csv")

data.covid <- read.csv(file.covid) %>% 
  filter(iso_code == "OWID_WRL") %>% 
  select(iso_code, location, date, total_vaccinations) %>% 
  as_tibble() %>% 
  mutate(date = as_date(date, format = "%d-%m-%y"),
         date.month = ceiling_date(date, "month") - 1) %>% 
  group_by(date.month) %>% 
  slice(n()) %>% 
  ungroup() %>% 
  filter(date == date.month) %>% 
  mutate(year = decimal_date(date),
         value = total_vaccinations/1E6) %>% 
  mutate(unit = "million doses",
         tech = "COVID vaccinations") %>% 
  select(year, value, unit, tech)

```

## Join datasets

```{r}
data.unconventional <- bind_rows(
  data.ww2aircraft,
  data.libertyships,
  data.usnuclearweapons,
  data.sovietnuclearweapons,
  data.nuclear,
  data.internet,
  data.chinahsr,
  data.shale,
  data.smartphones,
  data.ebikes,
  data.covid,
) %>% 
  group_by(tech) %>% 
  mutate(value.norm = value/max(value)) %>%
  ungroup() %>% 
  arrange(number, year)
```

## Estimate logistic growth rates

```{r}
# Loop over technologies and time slices
param.unconventional.fit <- NULL
techs <- data.unconventional %>% 
  filter(tech != "COVID vaccinations") %>% 
  pull(tech) %>% 
  unique()
for (i in techs) {
  # Filter technology
  data.tech <- data.unconventional %>%
    filter(tech == i)
  fit <- nls(value.norm ~ SSlogis(year, Asym, xmid, scal),
             data = data.tech)
  r <- exp(1/coef(fit)["scal"]) - 1  # Transform into annual growth rate
  # Temporary data to save
  temp <- tibble(
    tech = i,
    b = r
  )
  param.unconventional.fit <- bind_rows(param.unconventional.fit, temp)
}
```

## Re-parameterise growth distribution

```{r}
if (model.mode == "simulation"){
# New growth rate distribution
b.mean <- mean(param.unconventional.fit %>% pull(b))
b.sd <- sd(param.unconventional.fit %>% pull(b))
# New growth rate sample
b.sample <- rtruncnorm(n = n, a = 0.15, b = Inf, mean = b.mean, sd = b.sd)
data.input.unconv <- data.input
data.input.unconv$growth <- rep(b.sample, 2)  # For both EU and globally
# New input file
data.input.param.unconv <- data.input.param %>% 
  select(!c(growth.solar, growth.wind))
# New growth parameter values
data.input.param.unconv$growth.min <- bmin
data.input.param.unconv$growth.mean <- b.mean
data.input.param.unconv$growth.sd <- b.sd
data.input.param.unconv$growth.sample.mean <- mean(b.sample)
data.input.param.unconv$growth.sample.sd <- sd(b.sample)
data.input.param.unconv$growth.sample.q25 <- quantile(b.sample, 0.25)
data.input.param.unconv$growth.sample.q75 <- quantile(b.sample, 0.75)
# Save parameters and sample
saveRDS(data.input.param.unconv, "02_output_plots/04_unconventional_growth_parameters.rds")
saveRDS(data.input.unconv, "02_output_plots/05_unconventional_growth_sample.rds")
} else if (model.mode == "reproduction"){
  # Load parameters and sample
  data.input.param.unconv <- readRDS("02_output_plots/04_unconventional_growth_parameters.rds")
  data.input.unconv <- readRDS("02_output_plots/05_unconventional_growth_sample.rds")
}
```

## Call simulation

```{r}
if (model.mode == "simulation") {
  # Keep track of time
  start.time <- Sys.time()
  # Call diffusion model
  data.results.unconv <-
    calcTechnologyDiffusion(data.demand, data.input.unconv, 0.25)
  # Save simulation results
  saveRDS(data.results.unconv,
          "02_output_plots/06_unconventional_growth_results.rds")
  # Print duration of simulation
  end.time <- Sys.time()
  time.taken <- end.time - start.time
  print(time.taken)
  
} else if (model.mode == "reproduction") {
  data.results.unconv <-
    readRDS("02_output_plots/06_unconventional_growth_results.rds")
}
```

## FIGURE 5: Unconventional growth

```{r, warning = FALSE}
techs <- data.unconventional %>% 
  filter(tech != "COVID vaccinations") %>% 
  pull(tech) %>% 
  unique()
colors.tech <- pal_npg()(length(techs))
names(colors.tech) <- techs

top <- 1.1
mid <- 0.9
low <- 0.7

# Annotation labels
data.plot.labels <- tribble(
  ~year, ~ypos, ~color, ~label,
  1938, top, "WW2 US aircraft", "1: WWII US\naircraft",
  1946, mid, "WW2 US liberty ships", "2: WWII US\nliberty ships",
  1960, top, "US Nuclear weapons", "3: US nuclear\nweapons",
  1966, mid, "Soviet Nuclear weapons", "4: Soviet nuclear\nweapons",
  1986, top, "Nuclear France", "5: Nuclear power\nFrance",
  1998, low, "Internet host count", "6: Internet\nhosts",
  2005, top, "High-speed rail China", "7: High-speed\nrail China",
  2008, 1.27, "US shale gas", "8: US dry\nshale gas",
  2020, 1.27, "Smartphones globally", "9: Smartphones\nglobally",
  2021, top, "E-Bikes Europe", "10: E-Bikes\nEurope",
  2022.5, mid, "COVID vaccinations", "COVID-19\nvaccinations"
)

# PANEL a: Unconventional growth curves
p.curves <- ggplot() +
  geom_line(data = data.unconventional,
            mapping = aes(x = year, y = value.norm, color = tech)) + 
  scale_color_manual(name = NULL,
                     values = colors.tech) +
  scale_x_continuous(name = "",
                     breaks = seq(1940, 2020, 10),
                     limits = c(1938, 2032)) +
  scale_y_continuous(name = "") + 
  ggtitle("Normalised unconventional growth") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(
            face = "bold",
            size = font.size)) +
  geom_label(data = data.plot.labels,
            mapping = aes(x = year, y = ypos, label = label, color = color),
            size = 6 / .pt,
            hjust = 0,
            label.padding = unit(0.7, "mm")) +
  coord_cartesian(ylim = c(0, 1.2),clip = "off") +
  geom_line(mapping = aes(x = c(1938, 2024), y = 1),
            linetype = "dotted",
            alpha = 0.5)

growth.min <- 0.15
growth.max <- 4

# Sample
data.plot.sample <- data.input.unconv %>% 
  filter(region == "Global")  # The same anyway

# Density
data.plot.density <- tibble(
  growth = seq(growth.min, growth.max, length.out = 100),
  growth.dens = dtruncnorm(growth,
                           a = growth.min,
                           mean = data.input.param.unconv$growth.mean[[1]],
                           sd = data.input.param.unconv$growth.sd[[1]])
)

data.plot.growthrates <- param.unconventional.fit %>% 
  ungroup() %>% 
  mutate(number = 1:n()) %>% 
  arrange(b) %>% 
  mutate(order = 1:n()) %>% 
  mutate(ypos = case_when(order %% 4 == 1 ~ 0.45,
                          order %% 4 == 2 ~ 0.35,
                          order %% 4 == 3 ~ 0.25,
                          order %% 4 == 0 ~ 0.15))

bins <- 20

# PANEL b: Growth rates
p.growth <- ggplot() + 
  # Sample
  geom_histogram(data = data.plot.sample,
                 mapping = aes(x = 100*growth, y = 100*..density.., color = "Sample"),
                 binwidth = (100*growth.max - 100*growth.min)/bins,
                 boundary = 100*growth.min,
                 alpha = 0.3) +
  scale_color_manual(name = NULL, values = c("Sample" = "grey")) +
  # Distribution
  new_scale_color() +
  geom_line(data = data.plot.density,
            mapping = aes(x = 100*growth, y = growth.dens, color = "Density")) +
  scale_color_manual(name = NULL, values = c("Density" = "black")) +
  new_scale_color() + 
  # Vertical lines
  geom_vline(data = data.plot.growthrates,
             mapping = aes(xintercept = 100*b,
                           color = tech),
             linetype = "dashed") + 
  scale_color_manual(name = NULL,
                     values = colors.tech) +
  # Numbers in boxes
  geom_label(data = data.plot.growthrates,
             mapping = aes(x = b*100,
                           y = ypos,
                           label = number,
                           color =  tech),
             size = 6 /.pt,
             label.padding = unit(0.7, "mm")) +
  scale_x_continuous(name = "Emergence growth rate [%/yr]",
                     limits = c(0, 100*growth.max)) + 
  ylab("Probability density") +
  ggtitle("Distribution of unconventional growth rates") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(
          face = "bold",
          size = font.size))

# PANEL c + d: Unconventional electrolysis growth
colours <- pal_npg()(2)

p1 <-
  plotProbabilisticFeasibilitySpace(
    data.row = data.results.unconv %>%
      filter(region == "EU", anticipation == 5),
    data.targets = data.targets,
    colour = colours[2],
    marginal.year = 2030,
    title = "EU",
    ci.80 = T,
    zoom = F
  )

p2 <-
  plotProbabilisticFeasibilitySpace(
    data.row = data.results.unconv %>%
      filter(region == "Global", anticipation == 5),
    data.targets = data.targets,
    colour = colours[2],
    marginal.year = 2030,
    title = "Global",
    ci.80 = T,
    zoom = F
  )

# Arrange plots
p.row1 <- plot_grid(p.curves + theme(legend.position = "none"),
                    p.growth + theme(legend.position = "none"),
                    ncol = 2,
                    rel_widths = c(0.6, 0.4),
                    labels = c("a", "b"),
                    label_size = font.size)

p.row2 <- plot_grid(p1[[1]],
                    NULL,
                    p2[[1]],
                    NULL,
                    p1[[2]],
                    ncol = 5,
                    rel_widths = c(0.5, 0.025, 0.5, 0.05, 0.15),
                    labels = c("c", "", "d", "", ""),
                    label_size = font.size)

p <- plot_grid(p.row1,
               NULL,
               p.row2,
               ncol = 1,
               rel_heights = c(1, 0.05, 1))

print(p)
ggsave(paste0(path.output, "FIG05_unconventional_growth.png"),
       width = 18, height = 13, units = "cm", bg = "white")
ggsave(paste0(path.output, "FIG05_unconventional_growth.pdf"),
       width = 18, height = 13, units = "cm", bg = "white")
```

# Comparison of conventional and unconventional growth

## Final energy data

```{r, message = FALSE}
# EU data
# https://doi.org/10.1016/j.energy.2021.121908
# https://innopaths.eu/lcpp/#/
file.innopaths <- paste0(path.input, "innopaths_fe_se_h2_elec.csv")

data.fe.innopaths <- read.quitte(file.innopaths, sep = ";") %>% 
  filter(variable == "Final Energy",
         region %in% c("EU28", "UKI"),
         scenario %in% c("Efficient_1p5", "Incumb_1p5", "NewPl_1p5")) %>% 
  arrange(model, scenario, period, region) %>% 
  group_by(model, scenario, period) %>% 
  # EU27 ~= EU28 - UKI
  mutate(value = value - lead(value)) %>%
  filter(!is.na(value)) %>% 
  mutate(region = "EU27") %>%
  # Calculate median over all models and scenarios
  group_by(period, unit) %>% 
  summarise(median = median(value)) %>% 
  # PJ to TWh
  mutate(median.twh = median/3.6)

# Global data
# https://www.iea.org/data-and-statistics/data-product/net-zero-by-2050-scenario
file.ieanze <- paste0(path.input, "NZE2021_AnnexA.csv")

data.fe.nze <- read.csv(file.ieanze) %>% 
  as_tibble() %>% 
  filter(Category == "Energy",
         Product == "Total",
         Flow == "Total final consumption") %>% 
  # EJ to TWh
  mutate(value = 1E3/3.6*Value,
         unit = "TWh/yr") %>% 
  rename(period = Year) %>% 
  select(period, unit, value) %>% 
  complete(period = seq(2020, 2050, 5)) %>% 
  mutate(value = na.approx(value),
         unit = "TWh/yr")

```

## FIGURE 6: Distribution of breakthrough year and electrolysis capacity

```{r}
# EU breakthrough
p1.eu <- plotComparisonBreakthrough(regi = "EU",
                                    anticip = 5,
                                    title = NULL)

# EU 2030 capacity
p2.eu <- plotComparisonCapacity(
  regi = "EU",
  anticip = 5,
  title = "2030",
  final.energy = data.fe.innopaths %>%
    filter(period == 2030) %>%
    pull(median.twh),
  time.slice = 2030
)

# EU 2040 capacity
p3.eu <- plotComparisonCapacity(
  regi = "EU",
  anticip = 5,
  title = "2040",
  final.energy = data.fe.innopaths %>%
    filter(period == 2040) %>%
    pull(median.twh),
  time.slice = 2040
)

# Global breakthrough
p1.glob <- plotComparisonBreakthrough(regi = "Global",
                                      anticip = 5,
                                      title = NULL)

# Global 2030 capacity
p2.glob <- plotComparisonCapacity(
  regi = "Global",
  anticip = 5,
  title = "2030",
  final.energy = data.fe.nze %>%
    filter(period == 2030) %>%
    pull(value),
  time.slice = 2030
)

# Global 2040 capacity
p3.glob <- plotComparisonCapacity(
  regi = "Global",
  anticip = 5,
  title = "2040",
  final.energy = data.fe.nze %>%
    filter(period == 2040) %>%
    pull(value),
  time.slice = 2040
)

# Arranging plots
p.breakthrough <- plot_grid(p1.eu + theme(legend.position = "none"),
                            p1.glob + theme(legend.position = "none"),
                            ncol = 2,
                            labels = c("a", "b"),
                            label_size = font.size)

# Axes (lower panel)
axis.left <- ggdraw() +
  draw_label(
    "Electrolysis capacity [GW]",
    hjust = 0.5,
    size = font.size - 1,
    angle = 90
  )
axis.right <- ggdraw() + 
  draw_label(
    "Final energy share [%]",
    hjust = 0.5,
    size = font.size - 1,
    angle = 270
  )

# Lower panel
p.quantities <- plot_grid(axis.left,
                          p2.eu + theme(legend.position = "none"),
                          p3.eu + theme(legend.position = "none"),
                          p2.glob + theme(legend.position = "none"),
                          p3.glob + theme(legend.position = "none"),
                          axis.right,
                          ncol = 6,
                          rel_widths = c(0.1, 1, 1, 1, 1, 0.1),
                          labels = c("", "c", "d", "e", "f", ""),
                          label_size = font.size)

p.legend <- get_legend(p2.glob)

# Title top panel
title.top <- ggdraw() +
  draw_label(
    "Probability distribution of breakthrough year",
    hjust = 0.5,
    size = font.size,
    fontface = "bold"
  )

# Title lower panel
title.bottom <- ggdraw() +
  draw_label(
    "Probability distribution of electrolysis capacity",
    hjust = 0.5,
    size = font.size,
    fontface = "bold"
  )

# Title EU
title.eu <- ggdraw() +
  draw_label(
    "EU",
    hjust = 0.5,
    size = font.size + 1,
    fontface = "bold"
  )

# Title Global
title.glob <- ggdraw() +
  draw_label(
    "Global",
    hjust = 0.5,
    size = font.size + 1,
    fontface = "bold"
  )

# Combine region titles
title.reg <- plot_grid(title.eu, title.glob)

# Arrange full plot
p <- plot_grid(
  title.reg,
  title.top,
  p.breakthrough,
  NULL,
  title.bottom,
  p.quantities,
  p.legend,
  ncol = 1,
  rel_heights = c(0.1, 0.1, 1, 0.15, 0.1, 0.85, 0.2)
)

print(p)
ggsave(paste0(path.output, "FIG06_conventional_vs_unconventional.png"),
       width = 18, height = 12, units = "cm", bg = "white")
ggsave(paste0(path.output, "FIG06_conventional_vs_unconventional.pdf"),
       width = 18, height = 12, units = "cm")  # Leave out bg for editing
```

## Quantiles of breakthrough year

```{r, message = FALSE}
for (r in c("EU", "Global")){
  
  data.stat.conv <- data.results %>%
    filter(region == r, anticipation == 5) %>%
    select(!demand) %>%
    unnest(sensitivities) %>%
    unnest(results) %>%
    mutate(abs.growth = forecast - lag(forecast)) %>%
    group_by(sample) %>%
    filter(abs.growth == max(abs.growth)) %>%
    mutate(type = "Conventional growth")
  
  data.stat.unconv <- data.results.unconv %>%
    filter(region == r, anticipation == 5) %>%
    select(!demand) %>%
    unnest(sensitivities) %>%
    unnest(results) %>%
    mutate(abs.growth = forecast - lag(forecast)) %>%
    group_by(sample) %>%
    filter(abs.growth == max(abs.growth)) %>%
    mutate(type = "Unconventional growth")
  
  data.stat <- bind_rows(data.stat.conv, data.stat.unconv) %>%
    group_by(type) %>%
    summarise(quibble(year))
  
  print(data.stat)
}
```

## Quantiles of final energy share

```{r, message = FALSE}
t <- 2030

# EU
fe.eu <- data.fe.innopaths %>% 
  filter(period == t) %>% 
  pull(median.twh)

data.stat.conv <- data.results %>% 
  filter(region == "EU",
         anticipation == 5) %>% 
  unnest(sensitivities) %>%
  select(!demand) %>%
  unnest(results) %>%
  drop_na(forecast) %>%
  filter(year == t) %>% 
  mutate(share = 100 * forecast * flh * eff.fe * 1e-3 / fe.eu) %>% 
  mutate(type = "Conventional")

data.stat.unconv <- data.results.unconv %>% 
  filter(region == "EU",
         anticipation == 5) %>% 
  unnest(sensitivities) %>%
  select(!demand) %>%
  unnest(results) %>%
  drop_na(forecast) %>%
  filter(year == t) %>% 
  mutate(share = 100 * forecast * flh * eff.fe * 1e-3 / fe.eu) %>% 
  mutate(type = "Unconventional")  

data.stat <- bind_rows(data.stat.conv, data.stat.unconv) %>% 
  group_by(type) %>% 
  summarise(quibble(share))
print(data.stat)

# Global
fe.glob <- data.fe.nze %>% 
  filter(period == t) %>% 
  pull(value)

data.stat.conv <- data.results %>% 
  filter(region == "Global",
         anticipation == 5) %>% 
  unnest(sensitivities) %>%
  select(!demand) %>%
  unnest(results) %>%
  drop_na(forecast) %>%
  filter(year == t) %>% 
  mutate(share = 100 * forecast * flh * eff.fe * 1e-3 / fe.glob) %>% 
  mutate(type = "Conventional")

data.stat.unconv <- data.results.unconv %>% 
  filter(region == "Global",
         anticipation == 5) %>% 
  unnest(sensitivities) %>%
  select(!demand) %>%
  unnest(results) %>%
  drop_na(forecast) %>%
  filter(year == t) %>% 
  mutate(share = 100 * forecast * flh * eff.fe * 1e-3 / fe.glob) %>% 
  mutate(type = "Unconventional")  

data.stat <- bind_rows(data.stat.conv, data.stat.unconv) %>% 
  group_by(type) %>% 
  summarise(quibble(share))
print(data.stat)

```

## Probability of reaching 2030 targets

```{r}

t <- 2030

# EU
data.stat.conv <- data.results %>% 
  filter(region == "EU",
         anticipation == 5) %>% 
  unnest(sensitivities) %>%
  select(!demand) %>%
  unnest(results) %>%
  drop_na(forecast) %>%
  filter(year == t) %>% 
  mutate(type = "Conventional")

data.stat.unconv <- data.results.unconv %>% 
  filter(region == "EU",
         anticipation == 5) %>% 
  unnest(sensitivities) %>%
  select(!demand) %>%
  unnest(results) %>%
  drop_na(forecast) %>%
  filter(year == t) %>% 
  mutate(type = "Unconventional")  

data.stat <- bind_rows(data.stat.conv, data.stat.unconv) %>% 
  group_by(type) %>% 
  summarise(n_tot = n(),
            n_gt = sum(forecast > 100),
            p = n_gt/n_tot)
print(data.stat)

# Global
data.stat.conv <- data.results %>% 
  filter(region == "Global",
         anticipation == 5) %>% 
  unnest(sensitivities) %>%
  select(!demand) %>%
  unnest(results) %>%
  drop_na(forecast) %>%
  filter(year == t) %>% 
  mutate(type = "Conventional")

data.stat.unconv <- data.results.unconv %>% 
  filter(region == "Global",
         anticipation == 5) %>% 
  unnest(sensitivities) %>%
  select(!demand) %>%
  unnest(results) %>%
  drop_na(forecast) %>%
  filter(year == t) %>% 
  mutate(type = "Unconventional")  

data.stat <- bind_rows(data.stat.conv, data.stat.unconv) %>% 
  group_by(type) %>% 
  summarise(n_tot = n(),
            n_gt = sum(forecast > 850),
            p = n_gt/n_tot)
print(data.stat)

```


# Extended data figures

## EXT FIG 1: Wind and solar growth, implicit demand pull

```{r, message = FALSE, warning = FALSE}
# EU solar
p1 <-
  plotWindSolar(
    data.bp = data.bp,
    data.bp.fit = data.bp.fit,
    slice.len = 7,
    regi = "EU",
    tech = "solar",
    slice.st = 1997,
    color = color.pv
  )

# EU wind
p2 <-
  plotWindSolar(
    data.bp = data.bp,
    data.bp.fit = data.bp.fit,
    slice.len = 7,
    regi = "EU",
    tech = "wind",
    slice.st = 1997,
    color = color.wind
  )

# Global solar
p3 <-
  plotWindSolar(
    data.bp = data.bp,
    data.bp.fit = data.bp.fit,
    slice.len = 7,
    regi = "Global",
    tech = "solar",
    slice.st = 1996,
    color = color.pv
  )

# Global wind
p4 <-
  plotWindSolar(
    data.bp = data.bp,
    data.bp.fit = data.bp.fit,
    slice.len = 7,
    regi = "Global",
    tech = "wind",
    slice.st = 1995,
    color = color.wind
  )

# Arrange plots
p <- plot_grid(p1 + theme(legend.position = "none"),
               p3 + theme(legend.position = "none"),
               get_legend(p3),
               p2 + theme(legend.position = "none"),
               p4 + theme(legend.position = "none"),
               get_legend(p4),
               ncol = 3,
               labels = c("a", "b", "", "c", "d", ""),
               label_size = font.size,
               rel_widths = c(1, 1, 0.4))

print(p)
ggsave(paste0(path.output, "EXT_FIG01_wind_solar.jpg"),
       width = 18, height = 16, bg = "white", units = "cm")
```

## EXT FIG 2: Conventional growth distributions for different slice lengths

```{r}
plots <- NULL
for (r in 1:2){
  
  # Get values
  growth.min <- data.input.param$growth.min[[r]]
  growth.max <- 1.25  # Only for plotting
  
  data.plot.density <- NULL
  for (l in unique(data.bp.fit$slice.length)){
    
    temp <- tibble(
      growth = seq(growth.min, growth.max, length.out = 100),
      growth.dens = dtruncnorm(
        growth,
        a = growth.min,
        mean = data.bp.fit %>%
          filter(slice.length == l,
                 region == regions.bp$name[r]) %>%
          pull(b) %>%
          mean(),
        sd = data.bp.fit %>%
          filter(slice.length == l,
                 region == regions.bp$name[r]) %>%
          pull(b) %>%
          sd()
      ),
      slice.length = paste0(l, " years")
    ) %>% 
      order.levels(slice.length = c("5 years",
                                    "6 years",
                                    "7 years",
                                    "8 years",
                                    "9 years",
                                    "10 years"))
    
    data.plot.density <- bind_rows(data.plot.density, temp)
    
  }

  # Growth rate
  p.growth <- ggplot() + 
    # Distribution
    geom_line(data = data.plot.density,
              mapping = aes(x = 100*growth, y = growth.dens, color = slice.length)) +
    scale_color_npg(name = "Slice length") +
    geom_area(data = data.plot.density,
              mapping = aes(x = 100*growth, y = growth.dens, fill = slice.length),
              position = "identity",
              alpha = 0.05) +
    scale_fill_npg(name = "Slice length") +
    xlab("Emergence growth rate [%/yr]") + 
    ylab("Probability density [yr]") +
    ggtitle(regions.h2$name[[r]]) +
    theme(# Only line for y axis
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.title = element_text(size = 7,
                                    hjust = 0.5,
                                    face = "bold"))
  plots <- append(plots, list(p.growth))

}

# Arrange plots
p <- plot_grid(plots[[1]] + theme(legend.position = "none"),
               plots[[2]] + theme(legend.position = "none"),
               get_legend(plots[[2]]),
               nrow = 1,
               rel_widths = c(1, 1, 0.2),
               labels = c("a", "b"),
               label_size = 7)

print(p)
ggsave(paste0(path.output, "EXT_FIG02_growth_different_slice_length.jpg"),
       width = 16, height = 8, units = "cm", bg = "white")
```

## EXT FIG 3: Comparison with Gompertz model

```{r}
# Panel a and b: Line plot of mean to illustrate difference

data.input.example <- tibble(
  region = c("EU", "Global"),
  sample = 1,
  growth = data.input.param %>%
    pull(growth.sample.mean),
  start = data.input.param %>%
    pull(start.sample.mean),
  start.year = 2023
)

# Calculate logistic model
data.results.example <-
  calcTechnologyDiffusion(data.demand, data.input.example, 0.25)

# Calculate Gompertz model EU
data.plot.eu <- calcGompertzModel(data.results.example %>% 
                                            filter(region == "EU",
                                                   anticipation == 5),
                                          t.fix = 7) %>% 
  unnest(sensitivities) %>% 
  select(!demand) %>% 
  unnest(results)

# Calculate Gompertz model Global
data.plot.glob <- calcGompertzModel(data.results.example %>% 
                                            filter(region == "Global",
                                                   anticipation == 5),
                                            t.fix = 7) %>% 
  unnest(sensitivities) %>% 
  select(!demand) %>% 
  unnest(results)

# Panel a: EU line plot
p1 <- ggplot() +
  geom_line(
    data = data.plot.eu,
    mapping = aes(x = year, y = forecast, color = "Logistic")
  ) +
  geom_line(
    data = data.plot.eu,
    mapping = aes(x = year, y = forecast.gomp, color = "Gompertz")
  ) +
  scale_color_npg(name = "Diffusion path",
                  breaks = c("Logistic", "Gompertz")) +
  new_scale_color() +
  geom_line(
    data = data.plot.eu,
    mapping = aes(x = year, y = demand, color = "Logistic"),
    linetype = "dotted"
  ) +
  geom_line(
    data = data.plot.eu,
    mapping = aes(x = year, y = demand.full, color = "Gompertz"),
    linetype = "dotted"
  ) +
  scale_color_npg(name = "Demand pull",
                  breaks = c("Logistic", "Gompertz")) +
  xlab(NULL) +
  ylab("Electrolysis capacity [GW]") +
  ggtitle("EU, Model comparison") +
  theme(plot.title = element_text(
            face = "bold",
            size = font.size))
  
# Panel b: Global line plot
p2 <- ggplot() +
  geom_line(
    data = data.plot.glob,
    mapping = aes(x = year, y = forecast, color = "Logistic")
  ) +
  geom_line(
    data = data.plot.glob,
    mapping = aes(x = year, y = forecast.gomp, color = "Gompertz")
  ) +
  scale_color_npg(name = "Diffusion path",
                  breaks = c("Logistic", "Gompertz")) +
  new_scale_color() +
  geom_line(
    data = data.plot.glob,
    mapping = aes(x = year, y = demand, color = "Logistic"),
    linetype = "dotted"
  ) +
  geom_line(
    data = data.plot.glob,
    mapping = aes(x = year, y = demand.full, color = "Gompertz"),
    linetype = "dotted"
  ) +
  scale_color_npg(name = "Demand pull",
                  breaks = c("Logistic", "Gompertz")) +
  xlab(NULL) +
  ylab("Electrolysis capacity [GW]") +
  ggtitle("Global, Model comparison") +
  theme(plot.title = element_text(
            face = "bold",
            size = font.size))

# Panel c and d: Logistic model again for comparison
colours <- pal_npg()(2)

p3 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results %>%
    filter(region == "EU", anticipation == 5),
  data.targets = data.targets,
  colour = colours[2],
  marginal.year = 2038,
  title = "EU, Logistic model",
  zoom.panel = F
)

p4 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results %>%
    filter(region == "Global", anticipation == 5),
  data.targets = data.targets,
  colour = colours[2],
  marginal.year = 2045,
  title = "Global, Logistic model",
  zoom.panel = F
)

# Panel e and f: Gompertz model
# Calculate Gompertz model on condition that 7-year growth = emergence growth
# EU
data.results.gomp.eu <- calcGompertzModel(data.results %>% 
                                              filter(region == "EU",
                                                     anticipation == 5),
                                            t.fix = 7)

# Global
data.results.gomp.glob <- calcGompertzModel(data.results %>% 
                                              filter(region == "Global",
                                                     anticipation == 5),
                                            t.fix = 7)

p5 <- plotProbabilisticFeasibilitySpace(
  data.results.gomp.eu,
  data.targets = data.targets,
  colour = colours[1],
  marginal.year = 2050,
  title = "EU, Gompertz model",
  zoom.panel = F,
  gompertz = T)

p6 <- plotProbabilisticFeasibilitySpace(
  data.results.gomp.glob,
  data.targets = data.targets,
  colour = colours[1],
  marginal.year = 2050,
  title = "Global, Gompertz model",
  zoom.panel = F,
  gompertz = T)

# Arrange plots
l1 <- get_legend(p1)
p.row1 <- plot_grid(p1 + theme(legend.position = "none"),
                    NULL,
                    p2 + theme(legend.position = "none"),
                    NULL,
                    l1,
                    ncol = 5,
                    rel_widths = c(9,1,9,1,3),
                    labels = c("a","", "b", "",""),
                    label_size = font.size)
p.row2 <- plot_grid(p3[[1]],
                    p4[[1]],
                    p3[[2]],
                    ncol = 3,
                    labels = c("c", "d", ""),
                    label_size = font.size,
                    rel_widths = c(10,10,3))
p.row3 <- plot_grid(p5[[1]],
                    p6[[1]],
                    p5[[2]],
                    ncol = 3,
                    labels = c("e", "f", ""),
                    label_size = font.size,
                    rel_widths = c(10,10,3))
p <- plot_grid(p.row1,
               p.row2,
               p.row3,
               ncol = 1)
print(p)
ggsave(paste0(path.output, "EXT_FIG03_logistic_vs_gompertz.jpg"),
       width = 18, height = 18, units = "cm", bg = "white")
```

## EXT FIG 4: Conventional growth sensitivity analysis of anticipation

```{r}
p1 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results %>%
    filter(region == "EU", anticipation == 0),
  data.targets = data.targets,
  colour = colours[1],
  marginal.year = 2038,
  title = "EU, no anticipation",
  zoom.panel = F
)

p2 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results %>%
    filter(region == "Global", anticipation == 0),
  data.targets = data.targets,
  colour = colours[1],
  marginal.year = 2045,
  title = "Global, no anticipation",
  zoom.panel = F
)

p3 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results %>%
    filter(region == "EU", anticipation == 10),
  data.targets = data.targets,
  colour = colours[1],
  marginal.year = 2038,
  title = "EU, 10 years anticipation",
  zoom.panel = F
)

p4 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results %>%
    filter(region == "Global", anticipation == 10),
  data.targets = data.targets,
  colour = colours[1],
  marginal.year = 2045,
  title = "Global, 10 years anticipation",
  zoom.panel = F
)

p5 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results %>%
    filter(region == "EU", anticipation == 100),
  data.targets = data.targets,
  colour = colours[1],
  marginal.year = 2038,
  title = "EU, full anticipation",
  zoom.panel = F
)

p6 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results %>%
    filter(region == "Global", anticipation == 100),
  data.targets = data.targets,
  colour = colours[1],
  marginal.year = 2045,
  title = "Global, full anticipation",
  zoom.panel = F
)

p.legend <- p1[[2]]

p.plots <- plot_grid(p1[[1]], p2[[1]],
                     p3[[1]], p4[[1]],
                     p5[[1]], p6[[1]],
                     ncol = 2,
                     rel_widths = c(1, 1),
                     labels = "auto",
                     label_size = font.size)

p <- plot_grid(p.plots,
               NULL,
               p.legend,
               ncol = 3,
               rel_widths = c(1, 0.05, 0.15))

print(p)
ggsave(paste0(path.output, "EXT_FIG03_sensitivity_anticipation_conventional.jpg"),
       width = 18, height = 18, units = "cm", bg = "white")
```

## EXT FIG 5: Comparison with announcements (validation)

```{r}
# Conventional growth case colour
colour <- pal_npg()(2)[1]
col.quant <- c(
  "q95" = darken(colour, 0.1),
  "q50" = darken(colour, 0.5),
  "q5" = darken(colour, 0.1)
)

# EU
# Project announcements
data.plot.announcements <- data.h2 %>% 
  filter(year %in% seq(2023,2030),
         region %in% names(region.mapping.eu)) %>% 
  group_by(status, year) %>% 
  summarise(cumcap.sum = sum(cumcap.sum))

# Model results
data.plot.model <- data.results %>% 
  filter(region == "EU",
         anticipation == 5) %>% 
  unnest(sensitivities) %>%
  select(!demand) %>% 
  unnest(results) %>% 
  group_by(year) %>% 
  summarise(
    q5 = quantile(forecast, 0.05),
    q50 = quantile(forecast, 0.5),
    q95 = quantile(forecast, 0.95)) %>% 
  pivot_longer(cols = c("q5", "q50", "q95")) %>% 
  filter(year <= 2030)

p1 <- ggplot() + 
  # Projects
  geom_bar(data = data.plot.announcements,
           mapping = aes(x = year, y = cumcap.sum, fill = status),
           stat = "identity",
           alpha = 0.5) +
  scale_fill_npg(name = "Status") + 
  # Model
  geom_line(data = data.plot.model,
            mapping = aes(x = year, y = value, color = name)) + 
  scale_color_manual(name = "Diffusion model\npercentiles",
                     values = col.quant,
                     labels = c("q95" = "95 %",
                                "q50" = "50 % (median)",
                                "q5" = "5 %")) + 
  ggtitle("EU") + 
  xlab(NULL) +
  ylab("Electrolysis capacity [GW]") +
  theme(plot.title = element_text(
            face = "bold",
            size = font.size))

# Global
# Project announcements
data.plot.announcements <- data.h2 %>% 
  filter(year %in% seq(2023,2030)) %>%
  revalue.levels(region = region.mapping.eu) %>% 
  group_by(status, year) %>% 
  summarise(cumcap.sum = sum(cumcap.sum))

# Model results
data.plot.model <- data.results %>% 
  filter(region == "Global",
         anticipation == 5) %>% 
  unnest(sensitivities) %>%
  select(!demand) %>% 
  unnest(results) %>% 
  group_by(year) %>% 
  summarise(
    q5 = quantile(forecast, 0.05),
    q50 = quantile(forecast, 0.5),
    q95 = quantile(forecast, 0.95)) %>% 
  pivot_longer(cols = c("q5", "q50", "q95")) %>% 
  filter(year <= 2030)

p2 <- ggplot() + 
  # Projects
  geom_bar(data = data.plot.announcements,
           mapping = aes(x = year, y = cumcap.sum, fill = status),
           stat = "identity",
           alpha = 0.5) +
  scale_fill_npg(name = "Status") + 
  # Model
  geom_line(data = data.plot.model,
            mapping = aes(x = year, y = value, color = name)) + 
  scale_color_manual(name = "Diffusion model\npercentiles",
                     values = col.quant,
                     labels = c("q95" = "95 %",
                                "q50" = "50 % (median)",
                                "q5" = "5 %")) + 
  ggtitle("Global") + 
  xlab(NULL) +
  ylab("Electrolysis capacity [GW]") +
  theme(plot.title = element_text(
            face = "bold",
            size = font.size))

# All
p <- plot_grid(p1 + theme(legend.position = "none"),
               p2 + theme(legend.position = "none"),
               get_legend(p1),
               ncol = 3,
               rel_widths = c(1, 1, 0.4),
               labels = c("a", "b", ""),
               label_size = font.size)
print(p)
ggsave(paste0(path.output, "EXT_FIG05_validation_announcements.jpg"),
       width = 16, height = 8, units = "cm", bg = "white")
```

## EXT FIG 6: Unconventional growth sensitivity analysis of anticipation

```{r}

p1 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results.unconv %>%
    filter(region == "EU", anticipation == 0),
  data.targets = data.targets,
  colour = colours[2],
  marginal.year = 2030,
  title = "EU, no anticipation",
  zoom.panel = F
)

p2 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results.unconv %>%
    filter(region == "Global", anticipation == 0),
  data.targets = data.targets,
  colour = colours[2],
  marginal.year = 2030,
  title = "Global, no anticipation",
  zoom.panel = F
)

p3 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results.unconv %>%
    filter(region == "EU", anticipation == 10),
  data.targets = data.targets,
  colour = colours[2],
  marginal.year = 2030,
  title = "EU, 10 years anticipation",
  zoom.panel = F
)

p4 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results.unconv %>%
    filter(region == "Global", anticipation == 10),
  data.targets = data.targets,
  colour = colours[2],
  marginal.year = 2030,
  title = "Global, 10 years anticipation",
  zoom.panel = F
)

p5 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results.unconv %>%
    filter(region == "EU", anticipation == 100),
  data.targets = data.targets,
  colour = colours[2],
  marginal.year = 2030,
  title = "EU, full anticipation",
  zoom.panel = F
)

p6 <- plotProbabilisticFeasibilitySpace(
  data.row = data.results.unconv %>%
    filter(region == "Global", anticipation == 100),
  data.targets = data.targets,
  colour = colours[2],
  marginal.year = 2030,
  title = "Global, full anticipation",
  zoom.panel = F
)

p.legend <- p1[[2]]

p.plots <- plot_grid(p1[[1]], p2[[1]],
                     p3[[1]], p4[[1]],
                     p5[[1]], p6[[1]],
                     ncol = 2,
                     rel_widths = c(1, 1),
                     labels = "auto",
                     label_size = font.size)

p <- plot_grid(p.plots,
               NULL,
               p.legend,
               ncol = 3,
               rel_widths = c(1, 0.05, 0.15))

print(p)
ggsave(paste0(path.output, "EXT_FIG06_sensitivity_anticipation_unconventional.jpg"),
       width = 18, height = 18, units = "cm", bg = "white")

```

## EXT FIG 7: Electrolysis capacity distribution for additional years

```{r}

plots.eu <- NULL
plots.glob <- NULL

time.plot <- seq(2025, 2050, 5)

# Interpolate IEA NZE
data.fe.nze.interpolate <- data.fe.nze %>% 
  complete(period = seq(2020, 2050, 5)) %>% 
  mutate(value = na.approx(value),
         unit = "TWh/yr")

for (t in time.plot){
  
  p.eu <- plotComparisonCapacity(regi = "EU",
                                 anticip = 5,
                                 title = t,
                                 final.energy = data.fe.innopaths %>% 
                                   filter(period == t) %>% 
                                   pull(median.twh),
                                 time.slice = t)
  
  p.glob <- plotComparisonCapacity(regi = "Global",
                                   anticip = 5,
                                   title = t,
                                   final.energy = data.fe.nze.interpolate %>% 
                                     filter(period == t) %>% 
                                     pull(value),
                                   time.slice = t)
  
  plots.eu <- append(plots.eu, list(p.eu))
  
  plots.glob <- append(plots.glob, list(p.glob))
  
}

# Arrange plots
p.plots <- plot_grid(
  plots.eu[[1]] + theme(legend.position = "none"),
  plots.eu[[2]] + theme(legend.position = "none"),
  plots.glob[[1]] + theme(legend.position = "none"),
  plots.glob[[2]] + theme(legend.position = "none"),
  plots.eu[[3]] + theme(legend.position = "none"),
  plots.eu[[4]] + theme(legend.position = "none"),
  plots.glob[[3]] + theme(legend.position = "none"),
  plots.glob[[4]] + theme(legend.position = "none"),
  plots.eu[[5]] + theme(legend.position = "none"),
  plots.eu[[6]] + theme(legend.position = "none"),
  plots.glob[[5]] + theme(legend.position = "none"),
  plots.glob[[6]] + theme(legend.position = "none"),
  ncol = 4,
  labels = "auto",
  label_size = font.size
)

p.legend <- get_legend(plots.eu[[1]])

# Title EU
title.eu <- ggdraw() +
  draw_label(
    "EU",
    hjust = 0.5,
    size = font.size,
    fontface = "bold"
  )

# Title Global
title.glob <- ggdraw() +
  draw_label(
    "Global",
    hjust = 0.5,
    size = font.size,
    fontface = "bold"
  )

# Combine region titles
title.reg <- plot_grid(NULL,
                       title.eu,
                       title.glob,
                       NULL,
                       ncol = 4,
                       rel_widths = c(0.03, 1, 1, 0.03))

# Axes
axis.left <- ggdraw() +
  draw_label(
    "Electrolysis capacity [GW]",
    hjust = 0.5,
    size = font.size,
    angle = 90
  )
axis.right <- ggdraw() + 
  draw_label(
    "Final energy share [%]",
    hjust = 0.5,
    size = font.size,
    angle = 270
  )

# Add axes
p.with.axes <- plot_grid(
  axis.left,
  p.plots,
  axis.right,
  ncol = 3,
  rel_widths = c(0.03, 1, 0.03)
)

# Arrange full plot
p <- plot_grid(
  title.reg,
  p.with.axes,
  p.legend,
  ncol = 1,
  rel_heights = c(0.08, 1, 0.08)
)

print(p)
ggsave(paste0(path.output, "EXT_FIG07_capacity_distributions_2025-2050.jpg"),
       width = 18, height = 14, units = "cm", bg = "white")
```

## EXT FIG 8: Comparison with scenarios (IPCC AR6, IPCC SR1.5, others)

Download IPCC SR1.5 and IPCC AR6 data first.

```{r}

plot.years <- c(2030, 2040, 2050)

# IPCC AR6 metadata (vetted)
ipcc.ar6.meta <- read_xlsx("01_input_data/AR6_Scenarios_Database_metadata_indicators_v1.0.xlsx",
                           sheet = "meta_Ch3vetted_withclimate") %>% 
  rename(model = Model,
         scenario = Scenario,
         category = Category) %>% 
  select(model, scenario, category)

# IPCC SR1.5 metadata
ipcc.sr15.meta <- read_xlsx("01_input_data/sr15_metadata_indicators_r2.0.xlsx",
                            sheet = "meta") %>% 
  select(model, scenario, category)

# EU
# IPCC AR6 data
data.ipcc.ar6.eu <- read_csv("01_input_data/AR6_Scenarios_Database_R10_regions_v1.0.csv",
                              col_names = T,
                              show_col_types = F) %>% 
  filter(variable %in% c("Capacity|Hydrogen|Electricity", "Secondary Energy|Hydrogen|Electricity"),
         region == "R10EUROPE") %>% 
  select(model, scenario, region, variable, unit, `2020`:`2050`) %>% 
  full_join(ipcc.ar6.meta) %>% 
  # Filter non-vetted scenarios
  filter(!is.na(category)) %>% 
  # Filter for <2C (67%) scenarios
  filter(category %in% c("C1", "C2", "C3")) %>% 
  as.quitte() %>% 
  as_tibble() %>% 
  # Calculate electrolysis capacity from secondary energy
  calc_addVariable("`Capacity|Hydrogen|Electricity`" = 
                     paste("`Secondary Energy|Hydrogen|Electricity` * ", ej2gw),
                 units = "GW") %>% 
  filter(variable == "Capacity|Hydrogen|Electricity") %>% 
  # Remove calculated variable if it already existed
  group_by(model, scenario, region, variable, unit, period, category) %>% 
  slice(1) %>% 
  filter(period %in% seq(2020, 2050, 5)) %>% 
  # Linearly interpolate missing values in 5-year steps
  group_by(model, scenario, region, variable, unit, category) %>% 
  complete(period = seq(2020, 2050, 5)) %>% 
  mutate(value = na.approx(value)) %>% 
  # Exclude outlier
  filter(scenario != "Transport_Budg1100_ConvSyn") %>% 
  # For plotting
  ungroup() %>% 
  filter(period %in% plot.years) %>% 
  select(period, value) %>% 
  rename(year = period) %>% 
  mutate(x = 1,
         model = "IPCC AR6")

# INNOPATHS
data.innopaths.eu <- read_delim("01_input_data/innopaths_fe_se_h2_elec.csv",
                                delim = ";",
                                show_col_types = F) %>% 
  filter(variable == "Secondary Energy|Hydrogen|Electricity",
         region %in% c("EU28", "UKI"),
         scenario %in% c("Efficient_1p5", "Incumb_1p5", "NewPl_1p5")) %>% 
  as.quitte() %>% 
  as_tibble() %>% 
  filter(period %in% seq(2025, 2050, 5),
         model != "PRIMES") %>%  # No H2 in PRIMES
  # Calculate electrolysis capacity from secondary energy
  calc_addVariable("`Capacity|Hydrogen|Electricity`" = 
                     paste("`Secondary Energy|Hydrogen|Electricity` * ", 1E-3*ej2gw),
                 units = "GW",
                 only.new = T) %>% 
  # EU27 ~= EU28 - UKI
  group_by(model, scenario, period) %>% 
  mutate(value = value - lead(value)) %>%
  filter(!is.na(value)) %>% 
  mutate(region = "EU27") %>% 
  # For plotting
  filter(period %in% plot.years) %>% 
  ungroup() %>% 
  select(period, value) %>% 
  rename(year = period) %>% 
  mutate(x = 2,
         model = "INNOPATHS")

# Modelling results: Conventional growth
data.plot.conv.eu <- data.results %>% 
  filter(region == "EU",
         anticipation == 5) %>% 
  unnest(sensitivities) %>% 
  select(!demand) %>% 
  unnest(results) %>% 
  # For plotting
  filter(year %in% plot.years) %>% 
  select(year, forecast) %>% 
  rename(value = forecast) %>% 
  mutate(x = 3,
         model = "Conventional growth")

# Modelling results: Unconventional growth
data.plot.unconv.eu <- data.results.unconv %>% 
  filter(region == "EU",
         anticipation == 5) %>% 
  unnest(sensitivities) %>% 
  select(!demand) %>% 
  unnest(results) %>% 
  # For plotting
  filter(year %in% plot.years) %>% 
  select(year, forecast) %>% 
  rename(value = forecast) %>% 
  mutate(x = 4,
         model = "Unconventional growth")

# IAM scenarios
data.plot.eu.scenarios <- bind_rows(data.ipcc.ar6.eu, data.innopaths.eu) %>% 
  order.levels(model = c("IPCC AR6", "INNOPATHS"))

# Diffusion model
data.plot.eu.model <- bind_rows(data.plot.conv.eu, data.plot.unconv.eu)

# Others
data.plot.points.eu <- tribble(~x, ~year, ~value, ~source,
                               5, 2030, 100, "Policy targets",
                               5, 2040, NA_real_, "Policy targets",
                               5, 2050, 500, "Policy targets",
                               5, 2030, 190, "Hydrogen Europe\n(\"ambitious\")",
                               5, 2050, 643, "Hydrogen Europe\n(\"ambitious\")")

shapes.eu <- c("Policy targets" = 21,
               "Hydrogen Europe\n(\"ambitious\")" = 17)

colours <- pal_npg()(4)
colours.model <- c("Conventional growth" = colours[1],
                   "Unconventional growth" = colours[2])
colours.scenarios.eu <- c("IPCC AR6" = colours[3],
                          "INNOPATHS" = colours[4])

# EU plot
p1 <- ggplot() + 
  geom_boxplot(data = data.plot.eu.scenarios,
               mapping = aes(x = x, y = value, fill = factor(model)),
               outlier.alpha = 0.3,
               coef = 1E20) +
  scale_fill_manual(name = "IAM scenarios",
                    values = colours.scenarios.eu,
                    guide = guide_legend(order = 1)) + 
  new_scale_fill() + 
  geom_boxplot(data = data.plot.eu.model,
               mapping = aes(x = x, y = value, fill = factor(model)),
               outlier.alpha = 0.005,
               coef = 1E20) +
  scale_fill_manual(name = "Probabilistic diffusion model",
                    values = colours.model,
                    guide = guide_legend(order = 2)) + 
  geom_point(data = data.plot.points.eu,
             mapping = aes(x = x, y = value, shape = source)) +
  scale_shape_manual(name = "Others",
                     values = shapes.eu,
                     guide = guide_legend(order = 3)) +
  ggtitle("EU") + 
  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) + 
  ylab("Electrolysis capacity [GW]") +
  facet_wrap(~year, ncol = 1,
             scales = "free") +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        plot.title = element_text(size = 7,
                                  hjust = 0.5),
        strip.background=element_rect(fill = "white"),
        strip.text=element_text(face = "bold"))
  
# Global
# IPCC AR6 data
data.ipcc.ar6.glob <- read_csv("01_input_data/AR6_Scenarios_Database_World_v1.0.csv",
                               col_names = T,
                               show_col_types = F) %>% 
  filter(Variable %in% c("Capacity|Hydrogen|Electricity", "Secondary Energy|Hydrogen|Electricity")) %>%
  rename(model = Model,
         scenario = Scenario,
         region = Region,
         variable = Variable,
         unit = Unit) %>% 
  select(model, scenario, region, variable, unit, `2020`:`2050`) %>% 
  full_join(ipcc.ar6.meta) %>% 
  # Filter non-vetted scenarios
  filter(!is.na(category)) %>% 
  # Filter for <2C (67%) scenarios
  filter(category %in% c("C1", "C2", "C3")) %>% 
  as.quitte() %>% 
  as_tibble() %>% 
  filter(!is.na(value)) %>% 
  # Calculate electrolysis capacity from secondary energy
  calc_addVariable("`Capacity|Hydrogen|Electricity`" = 
                     paste("`Secondary Energy|Hydrogen|Electricity` * ", ej2gw),
                 units = "GW") %>% 
  filter(variable == "Capacity|Hydrogen|Electricity") %>% 
  # Remove calculated variable if it already existed
  group_by(model, scenario, region, variable, unit, period, category) %>% 
  slice(1) %>% 
  # Linearly interpolate missing values in 5-year steps
  group_by(model, scenario, region, variable, unit, category) %>% 
  complete(period = seq(2020, 2050, 5)) %>% 
  mutate(value = na.approx(value)) %>% 
  # Exclude outlier
  filter(scenario != "Transport_Budg1100_ConvSyn") %>% 
  # For plotting
  ungroup() %>% 
  filter(period %in% plot.years) %>% 
  select(period, value) %>% 
  rename(year = period) %>% 
  mutate(x = 2,
         model = "IPCC AR6")

# IPCC SR1.5
data.ipcc.sr15.glob <- read_xlsx("01_input_data/iamc15_scenario_data_world_r2.0.xlsx",
                                 sheet = "data") %>% 
  filter(Variable == "Secondary Energy|Hydrogen|Electricity") %>% 
  rename(model = Model,
         scenario = Scenario,
         region = Region,
         variable = Variable,
         unit = Unit) %>% 
  select(model, scenario, region, variable, unit, `2020`:`2050`) %>% 
  full_join(ipcc.sr15.meta) %>% 
  filter(category %in% c("Below 1.5C",
                         "1.5C low overshoot",
                         "1.5C high overshoot",
                         "Lower 2C")) %>% 
  as.quitte() %>% 
  as_tibble() %>%
  filter(period %in% seq(2020, 2050, 5),
         any(value > 0.1)) %>% 
  # Calculate electrolysis capacity from secondary energy
  calc_addVariable("`Capacity|Hydrogen|Electricity`" = 
                     paste("`Secondary Energy|Hydrogen|Electricity` * ", ej2gw),
                 units = "GW",
                 only.new = T) %>% 
  # Linearly interpolate missing values in 5-year steps
  group_by(model, scenario, region, variable, unit, category) %>% 
  complete(period = seq(2020, 2050, 5)) %>% 
  mutate(value = na.approx(value)) %>% 
  # For plotting
  ungroup() %>% 
  filter(period %in% plot.years) %>% 
  select(period, value) %>% 
  rename(year = period) %>% 
  mutate(x = 1,
         model = "IPCC SR1.5")

# Modelling results: Conventional growth
data.plot.conv.glob <- data.results %>% 
  filter(region == "Global",
         anticipation == 5) %>% 
  unnest(sensitivities) %>% 
  select(!demand) %>% 
  unnest(results) %>% 
  # For plotting
  filter(year %in% plot.years) %>% 
  select(year, forecast) %>% 
  rename(value = forecast) %>% 
  mutate(x = 3,
         model = "Conventional growth")

# Modelling results: Unconventional growth
data.plot.unconv.glob <- data.results.unconv %>% 
  filter(region == "Global",
         anticipation == 5) %>% 
  unnest(sensitivities) %>% 
  select(!demand) %>% 
  unnest(results) %>% 
  # For plotting
  filter(year %in% plot.years) %>% 
  select(year, forecast) %>% 
  rename(value = forecast) %>% 
  mutate(x = 4,
         model = "Unconventional growth")

# IAM scenarios
data.plot.glob.scenarios <- bind_rows(data.ipcc.ar6.glob, data.ipcc.sr15.glob) %>% 
  order.levels(model = c("IPCC SR1.5", "IPCC AR6"))

# Diffusion model
data.plot.glob.model <- bind_rows(data.plot.conv.glob, data.plot.unconv.glob)

# Others
data.plot.points.glob <- tribble(~x, ~year, ~value, ~source,
                                 5, 2030, 254, "Targets/Announcements",
                                 5, 2040, NA_real_, "Targets/Announcements",
                                 5.05, 2050, 3600, "Targets/Announcements",
                                 5, 2030, 600, "Hydrogen Council\n(\"green-only\")",
                                 5, 2050, 5500, "Hydrogen Council\n(\"green-only\")",
                                 5, 2030, 850, "IEA NZE",
                                 4.95, 2050, 3600, "IEA NZE",
                                 5, 2050, 5000, "IRENA 1.5C")

shapes.glob <- c("Targets/Announcements" = 21,
               "Hydrogen Council\n(\"green-only\")" = 17,
               "IEA NZE" = 18,
               "IRENA 1.5C" = 15)

colours <- pal_npg()(4)
colours.scenarios.glob <- c("IPCC AR6" = colours[3],
                            "IPCC SR1.5" = colours[4])

# Global plot
p2 <- ggplot() + 
  geom_boxplot(data = data.plot.glob.scenarios,
               mapping = aes(x = x, y = value, fill = factor(model)),
               outlier.alpha = 0.3,
               coef = 1E20) +
  scale_fill_manual(name = "IAM scenarios",
                    values = colours.scenarios.glob,
                    guide = guide_legend(order = 1)) + 
  new_scale_fill() + 
  geom_boxplot(data = data.plot.glob.model,
               mapping = aes(x = x, y = value, fill = factor(model)),
               outlier.alpha = 0.005,
               coef = 1E20) +
  scale_fill_manual(name = "Probabilistic diffusion model",
                    values = colours.model,
                    guide = guide_legend(order = 2)) + 
  geom_point(data = data.plot.points.glob,
             mapping = aes(x = x, y = value, shape = source)) +
  scale_shape_manual(name = "Others",
                     values = shapes.glob,
                     guide = guide_legend(order = 3)) +
  ggtitle("Global") + 
  scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) + 
  ylab("Electrolysis capacity [GW]") +
  facet_wrap(~year, ncol = 1,
             scales = "free") +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        plot.title = element_text(size = 7,
                                  hjust = 0.5),
        strip.background=element_rect(fill = "white"),
        strip.text=element_text(face = "bold"))

# TOTAL
p.plots <- plot_grid(p1 + theme(legend.position = "none"),
                     p2 + theme(legend.position = "none"),
                     ncol = 2)

p.leg <- plot_grid(get_legend(p1 + theme(legend.justification = "center")),
                   get_legend(p2 + theme(legend.justification = "center")),
                   ncol = 2)

p <- plot_grid(p.plots,
               p.leg,
               ncol = 1,
               rel_heights = c(1,0.2))

print(p)
ggsave(paste0(path.output, "EXT_FIG08_comparison_scenarios.jpg"),
       width = 18, height = 14, units = "cm", bg = "white")
```